% To be compiled with pdf LaTeX
% This file is to be included into master file via \input command
% Note that there is no \begin{document} \end{document} brackets!

\newpage
\section{Bibliography}
\label{sec:biblio}

\begin{thebibliography}{}

\bibitem{ConanCentroid}
O. Lardiere, R. Conan, R. Clare, C. Bradley, N. Hubin,
``Performance comparison of centroiding algorithms for laser guide star
wavefront sensing with extremely large telescopes,'' \emph{Appl. Opt.},
\textbf{49}, G78-G94 (2010).

\bibitem{WibergMaxGavel1}
D. M. Wiberg, C. E. Max, and D. T. Gavel, ``A spatial
non-dynamic LQG controller: Part 1, Application to
adaptive optics,'' in \emph{Proceedings of the 2004 IEEE
Conference on Decision and Control} (IEEE, 2004), pp.
3326 - 3332.

\bibitem{WibergMaxGavel2}
D. M. Wiberg, C. E. Max, and D. T. Gavel, ``A spatial
non-dynamic LQG controller: Part 2, Theory,'' in
\emph{Proceedings of the 2004 IEEE Conference on Decision and
Control} (IEEE, 2004), pp. 3333 - 3338.

\bibitem{WibergMaxGavel3}
D. M. Wiberg, C. E. Max, and D. T. Gavel, ``Geometric view of adaptive optics
control,'' JOSA A, Vol. 22, pp. 870 - 880 (2004).

\bibitem{Poyneer1}
L. A. Poyneer, B. Macintosh,
``Spatially filtered wave-front sensor for high-order adaptive optics,''
JOSA A, Vol. 21, Issue 5, pp. 810-819 (2004).

\bibitem{GMT-LTAO-CODR}
\emph{Giant Magellan Telescope. Laser Tomography Adaptive Optics. Conceptual
Design Review.} \\
Research School of Astronomy and Astrophysics, The College of
Physical and Mathematical Science, The Australian National University. \\
GMTO Document Number: GMT-01-00140. (2011)

\bibitem{ConanGMTmath}
Rodolphe Conan, \emph{GMT Mathematics Handbook.} Personal communication. \\
rconan@mso.anu.edu.au

\bibitem{UptonEllerbroek}
R. Upton, B. Ellerbroek,
``Gram–Schmidt orthogonalization of the Zernike polynomials
on apertures of arbitrary shape,''
Opt. Lett., Vol. 29, pp. 2840 -- 2842, (2004).

\bibitem{SchmidtProp}
J. D. Schmidt, \emph{Numerical Simulation of Optical Wave Propagation With examples in MATLAB}, SPIE (2010).

\bibitem{PalmaBagini}
C. Palma, V. Bagini, ``Extension of the Fresnel transform to ABCD
systems,'' JOSA A, Vol. 14, pp. 1774 -- 1779 (1997).


\end{thebibliography}
